### 2024-09-09 submission/supplement.Rmd
supplement figure legend

---
title: Supplement - Erosion of somatic tissue identity with loss of the X-linked intellectual disability factor KDM5C
output: pdf_document
geometry: margin = 1in
font-size: 11pt
header-includes:
   - \renewcommand{\figurename}{Supplementary Figure}
   - \usepackage{caption}
   - \captionsetup[figure]{font=footnotesize, labelfont=bf}
   - \usepackage{sectsty} 
   - \usepackage{helvet}
   - \renewcommand{\familydefault}{\sfdefault}
   - \usepackage{lineno}
   - \linenumbers
   - \renewcommand\linenumberfont{\normalfont\small}
   - \usepackage{setspace}
   - \usepackage{siunitx}
   - \spacing{1.5}
   - \pagenumbering{arabic}
   - \usepackage{indentfirst}

bibliography: references.bib
csl: cell-reports.csl
indent: true
editor_options:
  chunk_output_type: console
---

\newpage
\begin{figure}
  \centering
  \includegraphics{../submission/compiled_figs/Supplement_KDM5C_ChIPseq.pdf}
  \caption[KDM5C binds to a subset of germline RNA-seq differentially expressed genes]{\textbf{KDM5C binds to a subset of germline RNA-seq differentially expressed genes.} \textbf{A.} Bar graph of the number of germline-enriched DEGs with promoter KDM5C ChIP-seq peaks in wild-type EpiLCs (Top) and PNCs (Bottom). RNA-seq DEGs were classified as shared betwen EpiLCs and the brain (Common), unique to EpiLCs (EpiLC Only), or unique to one or multiple brain regions (Brain Only). \textbf{B.} Average bigwigs of two example RNA-seq DEGs dysregulated in EpiLCs but not the brain, \textit{Dazl} and \textit{Stra8}. Top is the RNA-seq tracks for wild-type (WT) and \textit{Kdm5c}-KO (5CKO) EpiLCs, bottom is the KDM5C ChIP-seq tracks, with the annotated transcription start site (TSS) for each gene. \textbf{C.} Same as B but for two example DEGs only dysregualted in the brain and not expressed in EpiLCs, \textit{Ccnb1ip1} and \textit{Spata18}. \textbf{D.} HOMER motif analysis of all KDM5C-unbound germline DEGs shows signficant enrichment of multiple RFX members and their X-box motif. \textbf{E.} KDM5C ChIP-seq shows no KDM5C accumulation at the \textit{Rfx2} promoter in EpiLCs or PNCs.}
  \label{figurelabel}
\end{figure}